\bf Fast, Scalable Phrase-Based SMT Decoding

### 2016-09-10 amta.2016/amta2016.tex
more

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\citep{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\citep{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure~\citep{junczys_tsd_2012b}. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 0.6 \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}

@@

Each sentence is decoded from start to finish using one worker thread. %The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.

As can be seen in Table~\ref{tab:optimizing-memory}, over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, dampening the scalability of the decoder. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
\begin{table}
\begin{center}
\small
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

@@

We shall investigate the effects of the following three stack configurations on model score and decoding speed:
\begin{enumerate}
  \item \vspace{-2 mm} coverage cardinality,
  \item \vspace{-2 mm} coverage,
  \item \vspace{-2 mm} coverage and end position of most recently translated source word.
\end{enumerate}
Coverage cardinality is the same as that in Moses and Joshua. Coverage configuration uses one stack per unique coverage vector.

@@

By grouping hypotheses according to coverage \emph{and} end position into groups we call \emph{'ministack'}, the distortion limit only needs to be checked for each group, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}. 
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
  \FORALL{$r$ in translation rules}
    \IF {$d(C_{h}, e_{hypo}, range_r)$}
      \FORALL{ $h$ in $ministack_{C, e}$} 
	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
	\STATE add $h'$ to $ministack_{C',e'}$
      \ENDFOR
    \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
\label{algo:Hypothesis Expansion with Sorted Stack}
\end{figure}

Also, stack pruning occurs on each hypothesis group independently, therefore, potentially affecting search errors and model scores.

%Table~\ref{tab:stack-configuration} and 
Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model, created by varying the cube-pruning pop-limit. The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits. It is also slower but the time/quality tradeoff is better overall with this stack configuration. Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}. This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
%\begin{table*}[t]
%\small
%\begin{center}
%\begin{tabular}{|r|r|r|r|r|r|r|} \hline
%Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
%		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
%100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
%500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
%1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
%5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
%10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
%15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
%20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
%\end{tabular}
%\end{center}
%\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
%\label{tab:stack-configuration}
%\end{table*}

>>>>

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\citep{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\citep{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure~\citep{junczys_tsd_2012b}. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model (5-gram) 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 0.6 \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}

@@

%Each sentence is decoded from start to finish using one worker thread. %The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.

Over 24\% of the Moses decoder running time is spent on memory management, Table~\ref{tab:optimizing-memory}. This increases to 39\% when 32 threads are used, dampening the scalability of the decoder. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
\begin{table}
\begin{center}
\small
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

@@

We investigated the effects of the following three stack configurations on model score and decoding speed:
\begin{enumerate}
  \item \vspace{-2 mm} coverage cardinality,
  \item \vspace{-2 mm} coverage,
  \item \vspace{-2 mm} coverage and end position of most recently translated source word.
\end{enumerate}
Coverage cardinality is the same as that in Moses and Joshua. Coverage configuration uses one stack per unique coverage vector.

@@

By grouping hypotheses according to coverage \emph{and} end position into groups we call \emph{'ministack'}, the distortion limit only needs to be checked for each group, Figure~\ref{algo:Hypothesis Expansion with Sorted Stack}. 
\begin{figure} [h]
\begin{algorithmic}
%\REQUIRE source sentence $s$, translation options
\FORALL{ $ministack_{C, e}$ in $stack_{|C|}$} 
  \FORALL{$r$ in translation rules}
    \IF {$d(C_{h}, e_{hypo}, range_r)$}
      \FORALL{ $h$ in $ministack_{C, e}$} 
	\STATE expand $h$ with $r$ $\rightarrow$ $h'$
	\STATE add $h'$ to $ministack_{C',e'}$
      \ENDFOR
    \ENDIF
  \ENDFOR % { translation options}
\ENDFOR % { hypothesis in $stack_i$} 
\end{algorithmic}
\caption{Hypothesis Expansion with Coverage \& End Position Stacks}
\label{algo:Hypothesis Expansion with Sorted Stack}
\end{figure}
However, stack pruning occurs on each hypothesis group independently, therefore, potentially affecting search errors and model scores.

%Table~\ref{tab:stack-configuration} and 
Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model, created by varying the cube-pruning pop-limit. None of the different stack configurations significantly outperform the others in either quality or decoding speed. However, the coverage \& end position produces slightly higher model scores at higher pop-limits, therefore, we continue to use this configuration throughout the rest of this paper.

%The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits. It is also slower but the time/quality tradeoff is better overall with this stack configuration. Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}. This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.

%\begin{table*}[t]
%\small
%\begin{center}
%\begin{tabular}{|r|r|r|r|r|r|r|} \hline
%Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
%		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
%100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
%500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
%1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
%5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
%10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
%15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
%20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
%\end{tabular}
%\end{center}
%\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
%\label{tab:stack-configuration}
%\end{table*}
